Relocating a cluster of earthquakes using a single station

### 2012-12-12 paper2/paper2.tex
incorporated most of Roel's comments - left a couple of the harder ones

Earthquake location is important for many applications. Locations
are required for:  magnitude
determination \citep{dr_Richter35a, dr_Gutenberg45a};
computing moment tensors \citep{dr_Sipkin02a};
seismological studies of the Earth's interior
\citep{dr_Spencer80a, dr_Kennett95a, dr_Curtis02a, dr_Kennett04a};
understanding strong motion and seismic attenuation
\citep{dr_Toro97a, dr_Campbell03a}
 and modeling earthquake hazard or risk
\citep{dr_Frankel00a, dr_Stirling02a, dr_Robinson06b}.

The accuracy required in earthquake location depends on the
application. For example, imaging the structure of a fracture system
from microseismicity requires greater detail than determining
whether a $M_w=7.5$ earthquake occurs offshore for tsunami warning.
 This paper focuses on reducing location uncertainty for
a cluster of events when they are recorded by a small
number of stations.

@@

Reported location uncertainties from relative techniques are  around
15 to 75\,m in local settings with good station coverage
\citep{dr_Ito85a, dr_Got94a, dr_Waldhauser99a,dr_Waldhauser08a}.
Here, `good coverage' implies multiple stations distributed across a
broad range of azimuthal directions. Relative location techniques
have been used to image active fault planes \citep{dr_Deichmann92a,
dr_Got94a, dr_Waldhauser99a, dr_Waldhauser02a, dr_Shearer05a};
studying rupture mechanics \citep{dr_Rubin99a, dr_Rubin02a};
interpreting magma movement in volcanoes \citep{dr_Fremont87a}; and
monitoring pumping induced seismicity \citep{dr_Lees98a, dr_Ake05a}.

@@

Coda refers to  later arriving waves in the seismogram that arise
from scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}. Coda
waves are ignored in most seismological applications due to the
complexity involved in modeling their generation. In this paper we
develop an approach for locating earthquakes using coda waves.

\citet{dr_Snieder05a} demonstrate that the coda of two earthquakes
can be used to estimate the separation between them. Their
technique, known as coda wave interferometry (CWI), is based on the
interference pattern between the coda waves. Unlike travel time
techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate settings. In this
paper we demonstrate how CWI separation estimates can be used to constrain
location with data from a single station.
Our technique can be used on coda waves alone or in combination with
travel times.

@@

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation}
where $\sigma_\tau$ is the standard deviation of the travel time
perturbation between the coda waves of two earthquakes, and $\alpha$
and $\beta$ are the near source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double couple.
For example, for two double couple sources displaced in the fault
plane,
\begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation}
whereas, for two point sources in a 2D acoustic medium
\begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation}
\citep{dr_Snieder05a}. In this paper we use equation
\ref{eq:LitReview-g-3d-dc} which assumes that the source mechanism
of both events are identical, an assumption likely to be true for
events in the same fault plane. \cite{dr_Robinson07c} explore the
impact of a change in mechanism.

The $\sigma_\tau$ in equation (\ref{eq:Theory-dcwi_g_sigmat}) is
related to $R_{max}$, the maximum of the cross correlation between
the coda of the two waveforms, and hence can be computed directly
from the recorded data. In this paper we use the autocorrelation
approach of \citet{dr_Robinson11a} when relating these parameters
and we apply a restricted time lag search when evaluating $R_{max}$.
These extensions to the original technique of \citet{dr_Snieder05a}
increase the range of applicability of CWI by 50\% (i.e. from 300 to
450\,m separation for 1–5\,Hz filtered coda waves).

\citet{dr_Robinson11a} show that coda waves provide only
probabilistic constraints on source separation and introduce a
Bayesian approach for describing the probability of true separation
given the CWI data. Their approach is summarised by
\begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the posterior function
indicating the probability of true separation $\widetilde{\delta}_t$ given the noisy CWI separation
estimates $\widetilde{\delta}_{CWIN}$; $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$
is the likelihood function (or forward model) giving the probability that the separation
estimates $\widetilde{\delta}_{CWIN}$ would be observed if the true separation was $\widetilde{\delta}_t$; and
$P(\widetilde{\delta}_t)$ is the prior PDF accounting for all a-priori information.
The tilde above the separation parameters in eq. (\ref{eq:-bayesian-noisyform})
indicates the use of a wavelength normalised separation parameter
\begin{equation}
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$)
with respect to dominant wavelength $\lambda_{d}$. In this paper we
consider a uniform prior over appropriate bounds which gives emphasis to the recorded data.
The procedure for computing the noisy likelihood
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by
\citet{dr_Robinson11a} and summarised in Appendix
\ref{sec-Appendix-noisylikelihood}. With these two pieces in place
we can compute the posterior
$P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ (or PDF) for the
separation between any pair of events directly from their coda
waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
to describe the location of multiple events whose maximum corresponding to
 the most probable combination of locations.
More importantly however, the PDF shall
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data.

For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian
formulation we write
\begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$,
$P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ and
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In equation
(\ref{eq-Bayes-three-locations}) $\mathbf{d}$ represents
observations that constrain the locations. They can be any
combination of travel times, geodetic information  or CWI
separations. For example, if coda waves are used we have
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 |
\widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3)$ where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the
wavelength normalised separation estimates. Alternatively, if we use
CWI and travel time data we may write $P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN},
\mathbf{\Delta}_{TT})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{\Delta}_{TT}|\mathbf{e}_1,
\mathbf{e}_2, \mathbf{e}_3)$ where $\mathbf{\Delta}_{TT}$ represent
travel time differences. In this derivation and in Section
\ref{sec:benchmarking} we focus on the constraints imposed by coda
waves, whereas in Section \ref{sec:CalaverasLoc-CWIandTT} we
demonstrate how CWI and travel time data can be combined.

@@

The event locations $\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n$
are defined by coordinates $\hat{x}$, $\hat{y}$ and $\hat{z}$
where the hat $\hat{.}$ indicates use of
a local coordinate system. We choose a local coordinate system
which removes ambiguity associated with transformations of the coordinate system.
In defining this coordinate system we fix the first event at the origin,
\begin{equation}
\label{eq:coord-const1}
\mathbf{e}_1 = (0,0,0),
\end{equation}
the second event on the positive $\hat{x}$-axis
\begin{equation}
\label{eq:coord-const2}
\mathbf{e}_2 = (\hat{x}_2, 0, 0), \hat{x}_2>0
\end{equation}
the third on the $\hat{x}-\hat{y}$ plane
\begin{equation}
\label{eq:coord-const3}
\mathbf{e}_3 = (\hat{x}_3,\hat{y}_3,0), \hat{y}_3>0
\end{equation}
and the fourth to
\begin{equation}
\label{eq:coord-const4}
\mathbf{e}_4 = (\hat{x}_4,\hat{y}_4,\hat{z}_4), \hat{z}_4>0.
\end{equation}
This coordinate system reduces translational (constraint \ref{eq:coord-const1})
and rotational (constraints \ref{eq:coord-const2} to \ref{eq:coord-const4}) non-uniqueness
without loss of generality. It is necessary to work with a local coordinate
system when using coda waves alone because the CWI technique constrains only event separation
between earthquakes. The inclusion of travel times in Section \ref{sec:CalaverasLoc-CWIandTT}
allows us to move to a global reference system.

@@

We design a 2D synthetic experiment (example 1) by randomly
selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq
\hat{x}, \hat{y} \leq 50$\,m. These are indicated with triangles in
Figure \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $v=3,300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using equation (\ref{eq:mu1}). This assumption ensures that the
sample mean of hypothetical separation estimates is consistent with
known CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimization procedure on the hypothetical CWI data yields the
circles in Figure \ref{fig-2D50eq-relocation_eg1}. The optimization
does not lead to the exact solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimization event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 2 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N= 2
\sigma_{1to5Hz}(\delta_t)$.
 Actual and optimization event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is simplistic because we fix
$\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase the
uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\sigma_{1to5Hz}(\delta_t)$, where
$\sigma_{1to5Hz}(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig. 4(b) of ][]{dr_Robinson11a}. Repeating the
optimization leads to the circles in Figure
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

@@

Three observations can be drawn from the error structure in Figures
\ref{fig-2D50eq-relocation_eg1} and \ref{fig-2D50eq-relocation_eg3}.
Firstly, the location errors depicted by gray bars increase between
examples 1 and 2 with the introduction of larger noise. Secondly,
the errors are larger for events at greater distances from the
center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those on the outside
are moderated by links from a limited number of directions. This
observation is analogous to problems associated with poor azimuthal
coverage in common triangulation problems such as individual
earthquake location from limited travel time data, or GPS
positioning with few satellites. Our third observation is that the
location errors are circular, despite our attempt to correct for
rotational non-uniqueness with the local coordinate system.

The local coordinate system works by constraining the location of
the first three earthquakes. Earthquake 1 is fixed at the origin,
earthquake 2 on the positive $\hat{x}$-axis and earthquake 3 has
$\hat{y}>0$. As the number of events increase the strength of these
constraints on later events weakens allowing small rotations of
events with respect to each other. That is, even though the
rotational freedom of the cluster is in principal removed by the
constraints imposed on the events (see equations
(\ref{eq:coord-const1}) to (\ref{eq:coord-const3})\footnote{equation
\ref{eq:coord-const4} is needed in 3D only}) we observe that in
practice the presence of noise allows the rotational non-uniqueness
to reappear. This is because the `easiest' way data noise can
propagate is into the direction which is least constrained by the
data. The same phenomena is observed in linear inversion where noise
creates large spurious model changes in directions of the
eigenvectors with the smallest singular values \citep{dr_Aster05a}.
Unfortunately, there is no obvious way to overcome this issue when
using coda waves alone since the constraints are based on
separation. Fortunately however, combining coda waves with
measurements of travel times alleviates this problem and facilitate
the removal of a local coordinate system altogether (see Section
\ref{sec:CalaverasLoc-CWIandTT}).

@@

Figures \ref{fig:optimisationresults-2Dsynth}(a) and (b) illustrates
the maximum $\Delta_{max}$ (top) and mean $\Delta_\mu$ (middle) of
the coordinate error as a function of increasing linkage. We show
the statistics for the `best' optimization solution (black) and for
the solution space when all 25 optimizations are considered (gray).
In the latter case the best solution is determined by the set of
event locations which lead to the smallest value of $L$. The error
in the best solution is consistent when 30\% or more of the branches
are used. The errors increase when only 10\% or 20\% of the
constraints are included. Interestingly, this breakdown around 20\%
to 30\% coincides with the point where the average number of
branches required to link each pair reaches 2 (see
Fig.\ref{fig:optimisationresults-2Dsynth} (bottom)). Since the
average number of branches can be computed in advance it can be used
as an indication of the inversion stability prior to optimization. A
higher breakdown is observed when all 25 solutions are considered
collectively. For example, the maximum coordinate error
$\Delta_{max}$ exceeds that for the best solution for linkage
$\leq$60\% confirming that the optimization is susceptible to local
minima and that a range of starting points should be considered.

Some optimizations fail to converge after 1200 iterations when the
linkage is 60\% or lower. All optimizations fail when then linkage
is 20\% or lower. Despite their failure to converge, the locations
at final iteration are close to the actual solution.

@@

When 70\% of the direct constraints are considered all optimization
results (gray) are consistent with the best solution (black). The
best solution constrains the event locations down to 30\% of the
direct links.

There is one notable difference between the 3D and 2D results. In 2D
the final iteration was close to the actual solution when the
optimization failed to converge. Conversely, in 3D the optimization
appears to converge to the correct solution or fail completely,
leading to a set of locations at final iteration which do not
resemble the actual solution. This is depicted in Figure
\ref{fig:optimisationresults-3Dsynth} by the absence of the gray and
black lines below 60\% and 30\% of the constraints, respectively.
The reason for this difference may be due to the increased number of
degrees of freedom in 3D requiring a greater number of iterations to
converge. Nevertheless, the accurate convergence of the best
solution for cases with 30\% linkage or higher is encouraging for
the potential of coda wave optimization to constrain earthquake
location.

>>>>

Earthquake location is important for many applications. Locations
are required for:  magnitude
determination \citep{dr_Richter35a, dr_Gutenberg45a};
computing moment tensors \citep{dr_Sipkin02a};
seismological studies of the Earth's interior
\citep{dr_Spencer80a, dr_Kennett95a, dr_Curtis02a, dr_Kennett04a};
understanding strong motion and seismic attenuation
\citep{dr_Toro97a, dr_Campbell03a}
 and modeling earthquake hazard or risk
\citep{dr_Frankel00a, dr_Stirling02a, dr_Robinson06b}.
The accuracy required in earthquake location depends on the
application. For example, imaging the structure of a fracture system
from microseismicity requires greater detail than determining
whether a $M_w=7.5$ earthquake occurs offshore for tsunami warning.
 This paper focuses on reducing location uncertainty for
a cluster of events when they are recorded by a small
number of stations.

@@

Reported location uncertainties from relative techniques are  around
15 to 75\,m in local settings with good station coverage
\citep{dr_Ito85a, dr_Got94a, dr_Waldhauser99a,dr_Waldhauser08a}.
Here, `good coverage' implies multiple stations distributed across a
broad range of azimuthal directions. Relative location techniques
have been used to image active fault planes \citep{dr_Deichmann92a,
dr_Got94a, dr_Waldhauser99a, dr_Waldhauser02a, dr_Shearer05a};
studying rupture mechanics \citep{dr_Rubin99a, dr_Rubin02a};
interpreting magma movement in volcanoes \citep{dr_Fremont87a}; and
monitoring pumping-induced seismicity \citep{dr_Lees98a, dr_Ake05a}.

@@

Coda refers to  later arriving waves in the seismogram that arise
from scattering  \citep{dr_Aki69a,dr_Snieder99a,dr_Snieder06a}. Coda
waves are ignored in most seismological applications due to the
complexity involved in modeling their generation. In this paper we
develop an approach for locating earthquakes using coda waves.
\citet{dr_Snieder05a} demonstrate that the coda of two earthquakes
can be used to estimate the separation between them. Their
technique, known as coda wave interferometry (CWI), is based on the
interference pattern between the coda waves. Unlike travel time
techniques, CWI
 does not require multiple stations or good azimuthal coverage.
In fact, it is possible to obtain estimates of separation using a single
station \citep{dr_Robinson07b}. This makes CWI particularly interesting
for regions where station density is low such as intraplate settings. In this
paper we demonstrate how CWI separation estimates can be used to constrain
location with data from a single station.
Our technique can be used on coda waves alone or in combination with
travel times.

@@

\citet{dr_Snieder05a} introduce a CWI based estimator of source
separation $\delta_{CWI}$ between two earthquakes
\begin{equation}
\label{eq:Theory-dcwi_g_sigmat}
\delta_{CWI}^2 = g(\alpha,\beta)\sigma_\tau^2,
\end{equation}
where $\sigma_\tau$ is the standard deviation of the travel time
perturbation between the coda waves of two earthquakes, and $\alpha$
and $\beta$ are the near-source $P$ and $S$ wave velocities,
respectively. The function $g$ depends on the type of excitation
(explosion, point force, double couple) and on the direction of
source displacement relative to the point force or double couple.
For example, for two double couple sources displaced in the fault
plane,
\begin{equation}
\label{eq:LitReview-g-3d-dc}
g(\alpha,\beta) = 7
\frac{\left(\frac{2}{\alpha^6}+\frac{3}{\beta^6}\right)}{\left(\frac{6}{\alpha^8}+\frac{7}{\beta^8}\right)},
\end{equation}
whereas, for two point sources in a 2D acoustic medium
\begin{equation}
\label{eq:LitReview-g-2d-ps} g(\alpha,\beta) = 2 \alpha^2
\end{equation}
\citep{dr_Snieder05a}. In this paper we use equation
\ref{eq:LitReview-g-3d-dc} which assumes that the source mechanism
of both events are identical, an assumption likely to be true for
events in the same fault plane. \cite{dr_Robinson07c} explore the
impact of a change in mechanism.

The $\sigma_\tau$ in equation (\ref{eq:Theory-dcwi_g_sigmat}) is
related to the maximum of the cross correlation between
the coda of the two waveforms, $R_{max}$, and hence can be computed directly
from the recorded data. The original formulation by \citet{dr_Sniedr05a}
used a Taylor series expansion to relate $\sigma_\tau$ and $R_{max}$ by
\begin{equation}
\label{eq:Rmax-disp-orig} R_{max}^{(t,t_w)} = 1 - \frac{1}{2}
\overline{\omega^2} \sigma_\tau^2,
\end{equation} 
In this paper we use the autocorrelation
approach of \citet{dr_Robinson11a} to relate the parameters directly
and we apply a restricted time lag search when evaluating $R_{max}$.
These extensions to the original technique of \citet{dr_Snieder05a}
increase the range of applicability of CWI by 50\% (i.e. from 300 to
450\,m separation for 1–5\,Hz filtered coda waves).

\citet{dr_Robinson11a} show that coda waves provide only
probabilistic constraints on source separation and introduce a
Bayesian approach for describing the probability of true separation
given the CWI data. Their approach is summarised by
\begin{equation}
\label{eq:-bayesian-noisyform}
P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN}) \propto P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)
\times P(\widetilde{\delta}_t)
\end{equation}
where  $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ is the posterior function
indicating the probability of true separation $\widetilde{\delta}_t$ given the noisy CWI separation
estimates $\widetilde{\delta}_{CWIN}$; $P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$
is the likelihood function (or forward model) giving the probability that the separation
estimates $\widetilde{\delta}_{CWIN}$ would be observed if the true separation was $\widetilde{\delta}_t$; and
$P(\widetilde{\delta}_t)$ is the prior PDF accounting for all a-priori information.
The tilde above the separation parameters in eq. (\ref{eq:-bayesian-noisyform})
indicates the use of a wavelength normalised separation parameter
\begin{equation}
\label{eq-normalisation-eqn}
\widetilde{\delta} = \frac{\delta}{\lambda_{d}},
\end{equation}
which measures separation ($\delta$ = $\delta_{CWIN}$ or $\delta_t$)
with respect to dominant wavelength $\lambda_{d}$. In this paper we
consider a uniform prior over appropriate bounds to ensure that the 
posterior function is dominated by the recorded data.
The procedure for computing the noisy likelihood
$P(\widetilde{\delta}_{CWIN}|\widetilde{\delta}_t)$ is derived by
\citet{dr_Robinson11a} and summarised in Appendix
\ref{sec-Appendix-noisylikelihood}. With these two pieces in place
we can compute the posterior
$P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$ (or PDF) for the
separation between any pair of events directly from their coda
waves.

We seek a probability density function (PDF) which links
 individual pairwise posteriors $P(\widetilde{\delta}_t|\widetilde{\delta}_{CWIN})$
to describe the location of multiple events whose maximum corresponds to
 the most probable combination of locations.
More importantly however, the PDF shall
quantify location uncertainty and provide information
on the degree to which individual events are constrained
 by the data. For convenience, we begin with three earthquakes having locations
$\mathbf{e}_1$, $\mathbf{e}_2$ and $\mathbf{e}_3$. Using a Bayesian
formulation we write
\begin{equation}
\label{eq-Bayes-three-locations}
P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3| \mathbf{d}) \propto P( \mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)
\times P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3),
\end{equation}
where $P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 | \mathbf{d})$,
$P(\mathbf{d}|\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ and
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3)$ are the posterior,
likelihood and prior functions, respectively. In equation
(\ref{eq-Bayes-three-locations}) $\mathbf{d}$ represents
observations that constrain the locations. They can be any
combination of travel times, geodetic information  or CWI
separations. For example, if coda waves are used we have
$P(\mathbf{e}_1, \mathbf{e}_2, \mathbf{e}_3 |
\widetilde{\mitbf{\delta}}_{CWIN})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN}|\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3)$ where $\widetilde{\mitbf{\delta}}_{CWIN}$ are the
wavelength normalised separation estimates. Alternatively, if we use
CWI and travel time data we may write $P(\mathbf{e}_1, \mathbf{e}_2,
\mathbf{e}_3 | \widetilde{\mitbf{\delta}}_{CWIN},
\mathbf{\Delta}_{TT})$ and
$P(\widetilde{\mitbf{\delta}}_{CWIN},\mathbf{\Delta}_{TT}|\mathbf{e}_1,
\mathbf{e}_2, \mathbf{e}_3)$ where $\mathbf{\Delta}_{TT}$ represent
travel time differences. In this derivation and in Section
\ref{sec:benchmarking} we focus on the constraints imposed by coda
waves, whereas in Section \ref{sec:CalaverasLoc-CWIandTT} we
demonstrate how CWI and travel time data can be combined.

@@

The event locations $\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n$
are defined by coordinates $\hat{x}$, $\hat{y}$ and $\hat{z}$
where the hat $\hat{.}$ indicates use of
a local coordinate system. We choose a local coordinate system
which removes ambiguity associated with transformations of the coordinate system. It is 
necessary to do this because the distances between events are invariant for rotations
and translations of the seismicity patterns and hence we know in advance that rotations 
and translations can not be resolved from CWI alone.  
In defining this coordinate system we fix the first event at the origin,
\begin{equation}
\label{eq:coord-const1}
\mathbf{e}_1 = (0,0,0),
\end{equation}
the second event on the positive $\hat{x}$-axis
\begin{equation}
\label{eq:coord-const2}
\mathbf{e}_2 = (\hat{x}_2, 0, 0), \hat{x}_2>0
\end{equation}
the third on the $\hat{x}-\hat{y}$ plane
\begin{equation}
\label{eq:coord-const3}
\mathbf{e}_3 = (\hat{x}_3,\hat{y}_3,0), \hat{y}_3>0
\end{equation}
and the fourth to
\begin{equation}
\label{eq:coord-const4}
\mathbf{e}_4 = (\hat{x}_4,\hat{y}_4,\hat{z}_4), \hat{z}_4>0.
\end{equation}
This coordinate system reduces translational (constraint \ref{eq:coord-const1})
and rotational (constraints \ref{eq:coord-const2} to \ref{eq:coord-const4}) non-uniqueness
without loss of generality. It is necessary to work with a local coordinate
system when using coda waves alone because the CWI technique constrains only event separation
between earthquakes. The inclusion of travel times in Section \ref{sec:CalaverasLoc-CWIandTT}
allows us to move to a global reference system.

@@

We design a 2D synthetic acoustic experiment (example 1) by randomly
selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq
\hat{x}, \hat{y} \leq 50$\,m. These are indicated with triangles in
Figure \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $\alpha=3,300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using equation (\ref{eq:mu1}). This assumption ensures that the
sample mean of hypothetical separation estimates is consistent with
known CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimization procedure on the hypothetical CWI data yields the
circles in Figure \ref{fig-2D50eq-relocation_eg1}. The optimization
does not lead to the exact solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimization event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 2 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N= 2
\epsilon(\delta_t)$.
 Actual and optimization event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is simplistic because we fix
$\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase the
uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\epsilon(\delta_t)$, where
$\epsilon(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig. 4(b) of ][]{dr_Robinson11a}. Repeating the
optimization leads to the circles in Figure
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

@@

Three observations can be drawn from the error structure in Figures
\ref{fig-2D50eq-relocation_eg1} and \ref{fig-2D50eq-relocation_eg3}.
Firstly, the location errors depicted by gray bars increase between
examples 1 and 2 with the introduction of larger noise. Secondly,
the errors are larger for events at greater distances from the
center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those on the outside
are moderated by links from a limited number of directions. This
observation is analogous to problems associated with poor azimuthal
coverage in common triangulation problems such as individual
earthquake location from limited travel time data, or GPS
positioning with few satellites. Our third observation is that the
location errors form a pattern of circular rotation, despite our attempt to correct for
rotational non-uniqueness with the local coordinate system.

The local coordinate system works by constraining the location of
the first three earthquakes. Earthquake 1 is fixed at the origin,
earthquake 2 on the positive $\hat{x}$-axis and earthquake 3 has
$\hat{y}>0$. As the number of events increase the strength of these
constraints on later events weakens allowing small rotations of
events with respect to each other. That is, even though the
rotational freedom of the cluster is in principal removed by the
constraints imposed on the events (see equations
(\ref{eq:coord-const1}) to (\ref{eq:coord-const3})\footnote{equation
\ref{eq:coord-const4} is needed in 3D only}) we observe that in
practice the presence of noise allows the rotational non-uniqueness
to reappear. This is because the `easiest' way data noise can
propagate is into the direction which is least constrained by the
data. The same phenomena is observed in linear inversion where noise
creates large spurious model changes in directions of the
eigenvectors with the smallest singular values \citep{dr_Aster05a}.
 Fortunately however, combining coda waves with
measurements of travel times alleviates this problem and facilitate
the removal of a local coordinate system altogether (see Section
\ref{sec:CalaverasLoc-CWIandTT}).

@@

Figures \ref{fig:optimisationresults-2Dsynth}(a) and (b) illustrates
the maximum $\Delta_{max}$ (top) and mean $\Delta_\mu$ (middle) of
the coordinate error as a function of increasing linkage. We show
the statistics for the `best' optimization solution (black) and for
the solution space when all 25 optimizations are considered (gray).
In the latter case the best solution is determined by the set of
event locations which lead to the smallest value of $L$. The error
in the best solution is consistent when 30\% or more of the branches
are used. The errors increase when only 10\% or 20\% of the
constraints are included. Interestingly, this breakdown around 20\%
to 30\% coincides with the point where the average number of
branches required to link each event pair reaches 2 (see
Fig.\ref{fig:optimisationresults-2Dsynth} (bottom)). Since the
average number of branches can be computed in advance it can be used
as an indication of the inversion stability prior to optimization. A
higher breakdown is observed when all 25 solutions are considered
collectively. For example, the maximum coordinate error
$\Delta_{max}$ exceeds that for the best solution for linkage
$\leq$60\% confirming that the optimization is susceptible to local
minima and that a range of starting points should be considered.
Some optimizations fail to converge after 1200 iterations when the
linkage is 60\% or lower. All optimizations fail when then linkage
is 20\% or lower. Despite their failure to converge, the locations
at final iteration are close to the actual solution.

@@

When 70\% of the direct constraints are considered all optimization
results (gray) are consistent with the best solution (black). The
best solution constrains the event locations down to 30\% of the
direct links. There is one notable difference between the 3D and 2D results. In 2D
the final iteration was close to the actual solution when the
optimization failed to converge. Conversely, in 3D the optimization
appears to converge to the correct solution or fail completely,
leading to a set of locations at final iteration which do not
resemble the actual solution. This is depicted in Figure
\ref{fig:optimisationresults-3Dsynth} by the absence of the gray and
black lines below 60\% and 30\% of the constraints, respectively.
The reason for this difference may be due to the increased number of
degrees of freedom in 3D requiring a greater number of iterations to
converge. Nevertheless, the accurate convergence of the best
solution for cases with 30\% linkage or higher is encouraging for
the potential of coda wave optimization to constrain earthquake
location.
